---
title: 'Solution: PCA explanation'
output:
  html_document:
    df_print: paged
  pdf_document: default
---


```{r include=FALSE}
# If you are using Windows and cannot proceed with the installation and package loading, you might need to run the following codes in this chunk:

#install.packages("installr",repos = "http://cran.us.r-project.org")
#library(installr)

#install.rtools()
#install.pandoc()
```


```{r setup, include=FALSE, message=FALSE, warning=FALSE}

#install.packages("knitr", repos = "http://cran.us.r-project.org")
#install.packages("dplyr", repos = "http://cran.us.r-project.org")
#install.packages("kableExtra", repos = "http://cran.us.r-project.org")
#install.packages("readxl", repos = "http://cran.us.r-project.org")
#install.packages("lme4", repos = "http://cran.us.r-project.org")
#install.packages("stats", repos = "http://cran.us.r-project.org")
#install.packages("corrplot", repos = "http://cran.us.r-project.org")
#install.packages("RColorBrewer", repos = "http://cran.us.r-project.org")
#install.packages("factoextra", repos = "http://cran.us.r-project.org")
#install.packages("attempt")
#install.packages("FeatureImpCluster")
#install.packages("flexclust")



library(knitr)
library(dplyr)
library(kableExtra)
library(readxl)
library(lme4)
library(stats)
library(ggplot2)
library(corrplot)
library(RColorBrewer)
library(factoextra)
library(FeatureImpCluster)
library(attempt)
library(flexclust)


knitr::opts_chunk$set(echo = TRUE)
set.seed(123)
```

<br/>
<br/>


## Preparation and Set Up

Before we start, let us do some preparation work to get ready!

We need to: 

* Create a folder and set working directory to that folder on your device,

* Download datasets needed for this tutorial onto your working directory, and

* Load datasets for analysis later as we proceed


The dataset of this tutorial is made available to you. The files is named *kmeans* and is stored in an *xlsx* format. 

To save this datset and for convenience of analysis for this tutorial, let us create a folder in your computer for and name the folder *kmeans*. And then let us download the two files to the folder (*kmeans*) that you just created. 

In order to load the datasets to R, we need to run only **part** of the following chunk of codes, depending on the **operating system** of your device. To avoid running the codes that are not applicable and receiving error messages, please add "#" before each line of non-applicable codes in the chunk below. For example, if you are using a Mac, you need to put "#" before code *setwd("H:/downloads/kmeans")*.



```{r warning=FALSE}
# Please run the following codes depending on the operating system on your laptop/PC. 

# if you are using iOS, and your *kmeans* folder is created under "Downloads" on your Mac: you will need to first set your working directory to *kmeans* folder:

setwd("~/Downloads/kmeans")

# if you are using Windows, and your *kmeans* folder is created in your H drive: you will need to first set your working directory to *kmeans* folder:

#setwd("H:/downloads/kmeans")

# load the dataset
data <- read.csv("user-clustering.csv")
#summary(data)
```

</br>
</br>

**Now we are ready to go!**

</br>
</br>


### Data Visualisations

Here is a brief pre-view of the first 10 rows of our dataset: 

```{r echo=FALSE}
data_view <- data [1:23195,1:12]
data_view<- data_view %>% distinct(uid,.keep_all = TRUE)

na.omit(data_view)

```

<br/>

Here is a brief description of each variable in the dataset:

<br/>
<br/>

We can further visualise the dataset and understand variable distributions across properties by first plotting histogram for each variable and then examining the correlation between each pair of variables. 

```{r}
# Plot histogram per variable

hist(data_view$attitude, border="light blue", col="light blue")
hist(data_view$follow, border="light blue", col="light blue")
hist(data_view$cinema, border="light blue", col="light blue")
hist(data_view$like, border="pink", col="pink")
hist(data_view$like_num, border="pink", col="pink")
hist(data_view$fans, border="pink", col="pink")

```

This table is to show the principal components to represent the dimensions.

```{r}

#Plot correlation analysis 

Clusterdata <- data_view[,c(3:4,6:12)]
Clusterdata <- scale(Clusterdata,center=T,scale=T)

#summary(Clusterdata)

#library("factoextra")
#data(Clusterdata)
Clusterdata.active <- Clusterdata
head(Clusterdata.active[, 1:7])

pca_data = prcomp(Clusterdata, center = TRUE, scale = TRUE)
summary(pca_data)
```


Showing the weight of each features to make up for dimensions.
```{r}
library(FactoMineR)
Clusterdata <- data_view[,c(3:4,6:12)]
Clusterdata <- scale(Clusterdata,center=T,scale=T)

Clusterdata.active <- Clusterdata
head(Clusterdata.active[, 1:9])


pca_clust <- PCA(Clusterdata, 
                scale.unit = FALSE,
                graph = F, 
                ncp = 9) 
summary(pca_clust)


```

PCA graph showing the potential distribution.
```{r}

plot.PCA(pca_clust, 
         choix = c("ind"),
         habillage = 1,
         select = "contrib5",
         invisible = "quali")

plot.PCA(pca_clust, choix = c("var"))
```
Plot the correlation of 2 dimensions.
```{r}
pca_dimdesc <- dimdesc(pca_clust)
pca_dimdesc$Dim.1
pca_dimdesc$Dim.2

```



<br/>


Have fun! 


<br/>
<br/>

